---
title: "Calculations"
author: "Joshua Paik and Igor Rivin"
date: "1/28/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE, warning = FALSE}
library(readr)
library(dplyr)
library(ggplot2)
library(MASS)
library(naniar)
df <- read_csv("./revised_data.csv")
df$age <- (2020- df$year)
df$citperyear <- df$citations/df$age
df$amscitperyear <- df$amscit/df$age
df$lettergroup <- as.factor(df$lettergroup)
df$gender <- as.factor(df$gender)
df$role <- as.factor(df$role)
df$highered <- as.factor(df$highered)
df$institution <- as.factor(df$institution)
df$research <- as.factor(df$research)
df$country <- as.factor(df$country)
df$security <- as.factor(df$security)
df$field <- as.factor(df$field)
df$simplefield <- as.factor(df$simplefield)
```

## Permutation Test Function

```{r}
# input data and the number of permutations
meanPermutation <- function(Data, n,n_a){
  output <- matrix(NA, ncol = 1, nrow = n)
  p<-n_a/length(Data)
  for(i in 1:n){
    #sample p% of the data
    X_index <- sample(1:length(Data), floor(p * length(Data)))
    Y_index <- setdiff(1:length(Data), X_index)
    X <- Data[X_index]
    Y <- Data[Y_index]
    #calculate the difference
    diff <- mean(X)-mean(Y)
    #store
    output[i, ] <- diff
  } 
  return(output)
}
```
## Age
```{r}
mean(filter(df, (lettergroup == "A Only"|lettergroup == "A and B"))$age, na.rm = TRUE)
mean(filter(df, (lettergroup == "B and C" | lettergroup == "B Only"|lettergroup == "A and B"))$age, na.rm = TRUE)
mean(filter(df, (lettergroup == "C Only"|lettergroup == "B and C"))$age, na.rm = TRUE)
median(filter(df, (lettergroup == "A Only"|lettergroup == "A and B"))$age, na.rm = TRUE)
median(filter(df, (lettergroup == "B and C"|lettergroup == "B Only"|lettergroup == "A and B"))$age, na.rm = TRUE)
median(filter(df, (lettergroup == "C Only"|lettergroup == "C and B"))$age, na.rm = TRUE)
```
```{r}
muA <- mean(filter(df, (lettergroup == "A Only"|lettergroup == "A and B"))$age, na.rm = TRUE)
muB <- mean(filter(df, (lettergroup == "B and C"|lettergroup == "B Only"|lettergroup == "A and B"))$age, na.rm = TRUE)
muC <- mean(filter(df, (lettergroup == "C Only"|lettergroup == "B and C"))$age, na.rm = TRUE)

val1 <- muA - muB
val2 <- muB - muC
val3 <- muA - muC

set.seed(0)

n_a <- length(na.omit(filter(df, (lettergroup == "A Only"|lettergroup == "A and B"))$age))
n_b <- length(na.omit(filter(df, (lettergroup == "B and C"|lettergroup == "B Only"|lettergroup == "A and B"))$age))
n_c <- length(na.omit(filter(df, (lettergroup == "C Only"|lettergroup == "B and C"))$age))

dist1 <- meanPermutation(na.omit(filter(df,(lettergroup == "A Only"|lettergroup == "A and B"|lettergroup == "B and C"|lettergroup == "B Only"))$age),10000,n_a)
dist2 <- meanPermutation(na.omit(filter(df,lettergroup == "B and C"|lettergroup == "B Only"|lettergroup == "A and B"|lettergroup == "C Only")$age),10000,n_b)
dist3 <- meanPermutation(na.omit(filter(df,lettergroup == "A Only"|lettergroup == "A and B"|lettergroup == "C Only"|lettergroup == "B and C")$age),10000,n_c)


sum(dist1 < val1)/length(dist1)
sum(dist2 < val2)/length(dist2)
sum(dist3 < val3)/length(dist3)
```


```{r, echo = FALSE, warning = FALSE}
plot1 <- ggplot(df, aes(x=age))+
  geom_histogram(data=filter(df, (lettergroup == "A Only"|lettergroup == "A and B")),fill = "red", alpha = 0.2, binwidth = 5) + 
  geom_histogram(data=filter(df, (lettergroup == "B and C"|lettergroup == "B Only"|lettergroup == "A and B")),fill = "blue", alpha = 0.2, binwidth = 5) +
  geom_histogram(data=filter(df, (lettergroup == "C Only"|lettergroup == "B and C")),fill = "green", alpha = 0.2, binwidth = 5) +
  # Place dotted lines for mean values
  geom_vline(xintercept = mean(filter(df, (lettergroup == "A Only"|lettergroup == "A and B"))$age, na.rm = TRUE), linetype="solid", color = "red", size=1.5) +
  geom_vline(xintercept = mean(filter(df, (lettergroup == "B Only"|lettergroup == "A and B"))$age, na.rm = TRUE), linetype="solid", color = "blue", size=1.5) +
  geom_vline(xintercept = mean(filter(df, (lettergroup == "C Only"|lettergroup == "B and C"))$age, na.rm = TRUE), linetype="solid", color = "green", size=1.5) +
  ggtitle("Histogram of the Age of Letter Signatories with Population Means Marked Vertically 
          Letter A = Red, Letter B = Blue, Letter C = Green")
plot1
```
## R1 Math Professors AMS Citations
```{r}
mean(filter(df, ((lettergroup == "A Only"|lettergroup == "A and B")&(role=="professor")&institution=="domesticr1"&field=="math"))$amscit, na.rm = TRUE)
mean(filter(df, ((lettergroup == "B and C"|lettergroup == "B Only"|lettergroup == "A and B")&(role=="professor")&institution=="domesticr1"&field=="math"))$amscit, na.rm = TRUE)
mean(filter(df, ((lettergroup == "C Only"|lettergroup == "B and C")&institution=="domesticr1"&(role=="professor")&field=="math"))$amscit, na.rm = TRUE)
median(filter(df, ((lettergroup == "A Only"|lettergroup == "A and B")&institution=="domesticr1"&(role=="professor")&field=="math"))$amscit, na.rm = TRUE)
median(filter(df, ((lettergroup == "B and C"|lettergroup == "B Only"|lettergroup == "A and B")&institution=="domesticr1"&(role=="professor")&field=="math"))$amscit, na.rm = TRUE)
median(filter(df, ((lettergroup == "C Only"|lettergroup == "B and C")&institution=="domesticr1"&(role=="professor")&field=="math"))$amscit, na.rm = TRUE)
```
```{r}
muA <- mean(filter(df, ((lettergroup == "A Only"|lettergroup == "A and B")&(role=="professor")&institution=="domesticr1"&field=="math"))$amscit, na.rm = TRUE)
muB <- mean(filter(df, ((lettergroup == "B and C"|lettergroup == "B Only"|lettergroup == "A and B")&(role=="professor")&institution=="domesticr1"&field=="math"))$amscit, na.rm = TRUE)
muC <- mean(filter(df, ((lettergroup == "C Only"|lettergroup == "B and C")&(role=="professor")&institution=="domesticr1"&field=="math"))$amscit, na.rm = TRUE)

val1 <- muA - muB
val2 <- muB - muC
val3 <- muA - muC

set.seed(0)


n_a <- length(na.omit(filter(df, (lettergroup == "A Only"|lettergroup == "A and B")&institution=="domesticr1"&field=="math"&role=="professor")$amscit))
n_b <- length(na.omit(filter(df, (lettergroup == "B and C"|lettergroup == "B Only"|lettergroup == "A and B")&institution=="domesticr1"&field=="math"&role=="professor")$amscit))
n_c <- length(na.omit(filter(df, (lettergroup == "C Only"|lettergroup == "B and C")&institution=="domesticr1"&field=="math"&role=="professor")$amscit))

dist1 <- meanPermutation(na.omit(filter(df,(lettergroup == "A Only"|lettergroup == "A and B"|lettergroup == "B and C"|lettergroup == "B Only"&institution=="domesticr1"&field=="math"&role=="professor"))$amscit),10000,n_a)
dist2 <- meanPermutation(na.omit(filter(df,lettergroup == "B and C"|lettergroup == "B Only"|lettergroup == "A and B"|lettergroup == "C Only"&institution=="domesticr1"&field=="math"&role=="professor")$amscit),10000,n_b)
dist3 <- meanPermutation(na.omit(filter(df,lettergroup == "A Only"|lettergroup == "A and B"|lettergroup == "C Only"|lettergroup == "B and C"&institution=="domesticr1"&field=="math"&role=="professor")$amscit),10000,n_c)


sum(dist1 < val1)/length(dist1)
sum(dist2 < val2)/length(dist2)
sum(dist3 < val3)/length(dist3)
```

```{r, warning=FALSE, echo = FALSE}
plot4 <- ggplot(df, aes(x=amscit))+
  geom_histogram(data=filter(df, ((lettergroup == "A Only"|lettergroup == "A and B")&institution=="domesticr1"&field=="math"&(role=="professor"))),fill = "red", alpha = 0.2, binwidth = 1000) + 
  geom_histogram(data=filter(df, ((lettergroup == "B and C"|lettergroup == "B Only"|lettergroup == "A and B")&institution=="domesticr1"&field=="math"&(role=="professor"))),fill = "blue", alpha = 0.2, binwidth = 1000) +
  geom_histogram(data=filter(df,( (lettergroup == "C Only"|lettergroup == "B and C")&institution=="domesticr1"&field=="math"&(role=="professor"))),fill = "green", alpha = 0.2, binwidth = 1000) +
  geom_vline(xintercept = mean(filter(df, ((lettergroup == "A Only"|lettergroup == "A and B")&institution=="domesticr1"&field=="math"&(role=="professor")))$amscit, na.rm = TRUE), linetype="solid", color = "red", size=1.5) +
  geom_vline(xintercept = mean(filter(df,( (lettergroup == "B and C"|lettergroup == "B Only"|lettergroup == "A and B")&institution=="domesticr1"&field=="math"&(role=="professor")))$amscit, na.rm = TRUE), linetype="solid", color = "blue", size=1.5) +
  geom_vline(xintercept = mean(filter(df, ((lettergroup == "C Only"|lettergroup == "B and C")&institution=="domesticr1"&field=="math"&(role=="professor")))$amscit, na.rm = TRUE), linetype="solid", color = "green", size=1.5) +
  ggtitle("Histogram of Math Sci Net Citation of R1 Math Professors
   with Population Means Marked Vertically 
   Letter A = Red, Letter B = Blue, Letter C = Green")
plot4
```

## AMS Citations per year R1 Math Professors Only

```{r , warning=FALSE, echo = FALSE}
plot9 <- ggplot(df, aes(x=amscitperyear))+
  geom_histogram(data=filter(df, (lettergroup == "A Only"|lettergroup == "A and B")&institution=="domesticr1"&field=="math"&(role=="professor")),fill = "red", alpha = 0.2, binwidth = 50) + 
  geom_histogram(data=filter(df, (lettergroup == "B and C"|lettergroup == "B Only"|lettergroup == "A and B")&institution=="domesticr1"&field=="math"&(role=="professor")),fill = "blue", alpha = 0.2, binwidth = 50) +
  geom_histogram(data=filter(df, (lettergroup == "C Only"|lettergroup == "B and C")&institution=="domesticr1"&field=="math"&(role=="professor")),fill = "green", alpha = 0.2, binwidth = 50) +
  geom_vline(xintercept = mean(filter(df, (lettergroup == "A Only"|lettergroup == "A and B")&institution=="domesticr1"&field=="math"&(role=="professor"))$amscitperyear, na.rm = TRUE), linetype="solid", color = "red", size=1.5) +
  geom_vline(xintercept = mean(filter(df, (lettergroup == "B and C"|lettergroup == "B Only"|lettergroup == "A and B")&institution=="domesticr1"&field=="math"&(role=="professor"))$amscitperyear, na.rm = TRUE), linetype="solid", color = "blue", size=1.5) +
  geom_vline(xintercept = mean(filter(df, (lettergroup == "C Only"|lettergroup == "B and C")&institution=="domesticr1"&field=="math"&(role=="professor"))$amscitperyear, na.rm = TRUE), linetype="solid", color = "green", size=1.5) +
  ggtitle("Histogram of MathSciNet Citations per Year of R1 Math Professors
   with Population Means Marked Vertically 
   Letter A = Red, Letter B = Blue, Letter C = Green")
plot9
```

```{r, echo=FALSE}
mean(filter(df, (lettergroup == "A Only"|lettergroup == "A and B")&institution=="domesticr1"&field=="math"&(role=="professor"))$amscitperyear, na.rm = TRUE)
mean(filter(df, (lettergroup == "B and C"|lettergroup == "B Only"|lettergroup == "A and B")&institution=="domesticr1"&field=="math"&(role=="professor"))$amscitperyear, na.rm = TRUE)
mean(filter(df, (lettergroup == "C Only"|lettergroup == "B and C")&institution=="domesticr1"&field=="math"&(role=="professor"))$amscitperyear, na.rm = TRUE)
median(filter(df, (lettergroup == "A Only"|lettergroup == "A and B")&institution=="domesticr1"&field=="math"&(role=="professor"))$amscitperyear, na.rm = TRUE)
median(filter(df, (lettergroup == "B and C"|lettergroup == "B Only"|lettergroup == "A and B")&institution=="domesticr1"&field=="math"&(role=="professor"))$amscitperyear, na.rm = TRUE)
median(filter(df, (lettergroup == "C Only"|lettergroup == "B and C")&institution=="domesticr1"&field=="math"&(role=="professor"))$amscitperyear, na.rm = TRUE)
```
```{r}
muA <- mean(filter(df, (lettergroup == "A Only"|lettergroup == "A and B")&institution=="domesticr1"&field=="math"&(role=="professor"))$amscitperyear, na.rm = TRUE)
muB <- mean(filter(df, (lettergroup == "B and C"|lettergroup == "B Only"|lettergroup == "A and B")&institution=="domesticr1"&field=="math"&(role=="professor"))$amscitperyear, na.rm = TRUE)
muC <- mean(filter(df, (lettergroup == "C Only"|lettergroup == "B and C")&institution=="domesticr1"&field=="math"&(role=="professor"))$amscitperyear, na.rm = TRUE)

val1 <- muA - muB
val2 <- muB - muC
val3 <- muA - muC

set.seed(0)


n_a <- length(na.omit(filter(df, (lettergroup == "A Only"|lettergroup == "A and B")&institution=="domesticr1"&field=="math"&(role=="professor"))$amscitperyear))
n_b <- length(na.omit(filter(df, (lettergroup == "B and C"|lettergroup == "B Only"|lettergroup == "A and B")&institution=="domesticr1"&field=="math"&(role=="professor"))$amscitperyear))
n_c <- length(na.omit(filter(df, (lettergroup == "C Only"|lettergroup == "B and C")&institution=="domesticr1"&field=="math"&(role=="professor"))$amscitperyear))


pop<- filter(df, ((role=="professor")&field=="math"&institution=="domesticr1"))

dist1 <- meanPermutation(na.omit(filter(pop,(lettergroup == "A Only"|lettergroup == "A and B"|lettergroup == "B and C"|lettergroup == "B Only"))$amscitperyear),10000,n_a)
dist2 <- meanPermutation(na.omit(filter(pop,lettergroup == "B and C"|lettergroup == "B Only"|lettergroup == "A and B"|lettergroup == "C Only")$amscitperyear),10000,n_b)
dist3 <- meanPermutation(na.omit(filter(pop,lettergroup == "A Only"|lettergroup == "A and B"|lettergroup == "C Only"|lettergroup == "B and C")$amscitperyear),10000,n_c)


sum(dist1 < val1)/length(dist1)
sum(dist1 < val2)/length(dist2)
sum(dist1 < val3)/length(dist3)
```

```{r}
n_a
```



## AMS R1 math profs 1.3
```{r}
df$amscitperyearpower <- df$amscit/(df$age^1.3)
mean(filter(df, (lettergroup == "A Only"|lettergroup == "A and B")&institution=="domesticr1"&field=="math"&role=="professor")$amscitperyearpower, na.rm = TRUE)
mean(filter(df, (lettergroup == "B and C"|lettergroup == "B Only"|lettergroup == "A and B")&institution=="domesticr1"&field=="math"&role=="professor")$amscitperyearpower, na.rm = TRUE)
mean(filter(df, (lettergroup == "C Only"|lettergroup == "B and C")&institution=="domesticr1"&field=="math"&role=="professor")$amscitperyearpower, na.rm = TRUE)
median(filter(df, (lettergroup == "A Only"|lettergroup == "A and B")&institution=="domesticr1"&field=="math"&role=="professor")$amscitperyearpower, na.rm = TRUE)
median(filter(df, (lettergroup == "B and C"|lettergroup == "B Only"|lettergroup == "A and B")&institution=="domesticr1"&field=="math"&role=="professor")$amscitperyearpower, na.rm = TRUE)
median(filter(df, (lettergroup == "C Only"|lettergroup == "B and C")&institution=="domesticr1"&field=="math"&role=="professor")$amscitperyearpower, na.rm = TRUE)
```

```{r}
muA <- mean(filter(df, (lettergroup == "A Only"|lettergroup == "A and B")&institution=="domesticr1"&field=="math"&role=="professor")$amscitperyearpower, na.rm = TRUE)
muB <- mean(filter(df, (lettergroup == "B and C"|lettergroup == "B Only"|lettergroup == "A and B")&institution=="domesticr1"&field=="math"&role=="professor")$amscitperyearpower, na.rm = TRUE)
muC <- mean(filter(df, (lettergroup == "C Only"|lettergroup == "B and C")&institution=="domesticr1"&field=="math"&role=="professor")$amscitperyearpower, na.rm = TRUE)

val1 <- muA - muB
val2 <- muB - muC
val3 <- muA - muC

set.seed(0)
n_a <- length(na.omit(filter(df, (lettergroup == "A Only"|lettergroup == "A and B")&institution=="domesticr1"&field=="math"&(role=="professor"))$amscitperyearpower))
n_b <- length(na.omit(filter(df, (lettergroup == "B and C"|lettergroup == "B Only"|lettergroup == "A and B")&institution=="domesticr1"&field=="math"&(role=="professor"))$amscitperyearpower))
n_c <- length(na.omit(filter(df, (lettergroup == "C Only"|lettergroup == "B and C")&institution=="domesticr1"&field=="math"&(role=="professor"))$amscitperyearpower))


pop<- filter(df, ((role=="professor")&field=="math"&institution=="domesticr1"))

dist1 <- meanPermutation(na.omit(filter(pop,(lettergroup == "A Only"|lettergroup == "A and B"|lettergroup == "B and C"|lettergroup == "B Only"))$amscitperyearpower),10000,n_a)
dist2 <- meanPermutation(na.omit(filter(pop,lettergroup == "B and C"|lettergroup == "B Only"|lettergroup == "A and B"|lettergroup == "C Only")$amscitperyearpower),10000,n_b)
dist3 <- meanPermutation(na.omit(filter(pop,lettergroup == "A Only"|lettergroup == "A and B"|lettergroup == "C Only"|lettergroup == "B and C")$amscitperyearpower),10000,n_c)


sum(dist1 < val1)/length(dist1)
sum(dist2 < val2)/length(dist2)
sum(dist3 < val3)/length(dist3)
```

```{r , warning=FALSE, echo = FALSE}
plot9 <- ggplot(df, aes(x=amscitperyearpower))+
  geom_histogram(data=filter(df, (lettergroup == "A Only"|lettergroup == "A and B")&institution=="domesticr1"&field=="math"&role=="professor"),fill = "red", alpha = 0.2, binwidth = 5) + 
  geom_histogram(data=filter(df, (lettergroup == "B and C"|lettergroup == "B Only"|lettergroup == "A and B")&institution=="domesticr1"&field=="math"&role=="professor"),fill = "blue", alpha = 0.2, binwidth = 5) +
  geom_histogram(data=filter(df, (lettergroup == "C Only"|lettergroup == "B and C")&institution=="domesticr1"&field=="math"&role=="professor"),fill = "green", alpha = 0.2, binwidth = 5) +
  geom_vline(xintercept = mean(filter(df, (lettergroup == "A Only"|lettergroup == "A and B")&institution=="domesticr1"&field=="math"&role=="professor")$amscitperyearpower, na.rm = TRUE), linetype="solid", color = "red", size=1.5) +
  geom_vline(xintercept = mean(filter(df, (lettergroup == "B and C"|lettergroup == "B Only"|lettergroup == "A and B")&institution=="domesticr1"&field=="math"&role=="professor")$amscitperyearpower, na.rm = TRUE), linetype="solid", color = "blue", size=1.5) +
  geom_vline(xintercept = mean(filter(df, (lettergroup == "C Only"|lettergroup == "B and C")&institution=="domesticr1"&field=="math"&role=="professor")$amscitperyearpower, na.rm = TRUE), linetype="solid", color = "green", size=1.5) +
  ggtitle("Histogram of MathSciNet Citations per Year^1.3 of R1 Math Professors
   with Population Means Marked Vertically 
   Letter A = Red, Letter B = Blue, Letter C = Green")
plot9
```

## h-index
```{r}
mean(filter(df, (lettergroup == "A Only"|lettergroup == "A and B"))$hindex, na.rm = TRUE)
mean(filter(df, (lettergroup == "B and C"|lettergroup == "B Only"|lettergroup == "A and B"))$hindex, na.rm = TRUE)
mean(filter(df, (lettergroup == "C Only"|lettergroup == "B and C"))$hindex, na.rm = TRUE)
median(filter(df, (lettergroup == "A Only"|lettergroup == "A and B"))$hindex, na.rm = TRUE)
median(filter(df, (lettergroup == "B and C"|lettergroup == "B Only"|lettergroup == "A and B"))$hindex, na.rm = TRUE)
median(filter(df, (lettergroup == "C Only"|lettergroup == "B and C"))$hindex, na.rm = TRUE)
```
```{r}
muA <- mean(filter(df, (lettergroup == "A Only"|lettergroup == "A and B"))$hindex, na.rm = TRUE)
muB <- mean(filter(df, (lettergroup == "B and C"|lettergroup == "B Only"|lettergroup == "A and B"))$hindex, na.rm = TRUE)
muC <- mean(filter(df, (lettergroup == "C Only"|lettergroup == "B and C"))$hindex, na.rm = TRUE)

val1 <- muA - muB
val2 <- muB - muC
val3 <- muA - muC


n_a <- length(na.omit(filter(df, (lettergroup == "A Only"|lettergroup == "A and B"))$hindex))
n_b <- length(na.omit(filter(df, (lettergroup == "B and C"|lettergroup == "B Only"|lettergroup == "A and B"))$hindex))
n_c <- length(na.omit(filter(df, (lettergroup == "C Only"|lettergroup == "B and C"))$hindex))

dist1 <- meanPermutation(na.omit(filter(df,(lettergroup == "A Only"|lettergroup == "A and B"|lettergroup == "B and C"|lettergroup == "B Only"))$hindex),10000,n_a)
dist2 <- meanPermutation(na.omit(filter(df,lettergroup == "B and C"|lettergroup == "B Only"|lettergroup == "A and B"|lettergroup == "C Only")$hindex),10000,n_b)
dist3 <- meanPermutation(na.omit(filter(df,lettergroup == "A Only"|lettergroup == "A and B"|lettergroup == "C Only"|lettergroup == "B and C")$hindex),10000,n_c)


sum(dist1 < val1)/length(dist1)
sum(dist2 < val2)/length(dist2)
sum(dist3 < val3)/length(dist3)
```

## Google Scholar Citations

```{r}
mean(filter(df, (lettergroup == "A Only"|lettergroup == "A and B"))$citations, na.rm = TRUE)
mean(filter(df, (lettergroup == "B and C"|lettergroup == "B Only"|lettergroup == "A and B"))$citations, na.rm = TRUE)
mean(filter(df, (lettergroup == "C Only"|lettergroup == "B and C"))$citations, na.rm = TRUE)
median(filter(df, (lettergroup == "A Only"|lettergroup == "A and B"))$citations, na.rm = TRUE)
median(filter(df, (lettergroup == "B and C"|lettergroup == "B Only"|lettergroup == "A and B"))$citations, na.rm = TRUE)
median(filter(df, (lettergroup == "C Only"|lettergroup == "B and C"))$citations, na.rm = TRUE)
```

```{r}
muA <- mean(filter(df, (lettergroup == "A Only"|lettergroup == "A and B"))$citations, na.rm = TRUE)
muB <- mean(filter(df, (lettergroup == "B Only"|lettergroup == "A and B"|lettergroup == "B and C"))$citations, na.rm = TRUE)
muC <- mean(filter(df, (lettergroup == "C Only"|lettergroup == "B and C"))$citations, na.rm = TRUE)

val1 <- muA - muB
val2 <- muB - muC
val3 <- muA - muC

set.seed(0)

n_a <- length(na.omit(filter(df, (lettergroup == "A Only"|lettergroup == "A and B"))$citations))
n_b <- length(na.omit(filter(df, (lettergroup == "B and C"|lettergroup == "B Only"|lettergroup == "A and B"))$citations))
n_c <- length(na.omit(filter(df, (lettergroup == "C Only"|lettergroup == "B and C"))$citations))

dist1 <- meanPermutation(na.omit(filter(df,(lettergroup == "A Only"|lettergroup == "A and B"|lettergroup == "B and C"|lettergroup == "B Only"))$citations),10000,n_a)
dist2 <- meanPermutation(na.omit(filter(df,lettergroup == "B and C"|lettergroup == "B Only"|lettergroup == "A and B"|lettergroup == "C Only")$citations),10000,n_b)
dist3 <- meanPermutation(na.omit(filter(df,lettergroup == "A Only"|lettergroup == "A and B"|lettergroup == "C Only"|lettergroup == "B and C")$citations),10000,n_c)


sum(dist1 < val1)/length(dist1)
sum(dist2 < val2)/length(dist2)
sum(dist3 < val3)/length(dist3)
```

## Google Scholar Citations Only Math R1 Professors

```{r}
mean(filter(df, ((lettergroup == "A Only"|lettergroup == "A and B")&(role=="professor")&field=="math"&institution=="domesticr1"))$citations, na.rm = TRUE)
mean(filter(df, ((lettergroup == "B and C"|lettergroup == "B Only"|lettergroup == "A and B")&(role=="professor")&field=="math"&institution=="domesticr1"))$citations, na.rm = TRUE)
mean(filter(df, ((lettergroup == "C Only"|lettergroup == "B and C")&(role=="professor")&field=="math"&institution=="domesticr1"))$citations, na.rm = TRUE)
median(filter(df, ((lettergroup == "A Only"|lettergroup == "A and B")&(role=="professor")&field=="math"&institution=="domesticr1"))$citations, na.rm = TRUE)
median(filter(df, ((lettergroup == "B Only"|lettergroup == "A and B")&(role=="professor")&field=="math"&institution=="domesticr1"))$citations, na.rm = TRUE)
median(filter(df, ((lettergroup == "C Only"|lettergroup == "B and C")&(role=="professor")&field=="math"&institution=="domesticr1"))$citations, na.rm = TRUE)
```

```{r}
muA <- mean(filter(df, ((lettergroup == "A Only"|lettergroup == "A and B")&(role=="professor"&field=="math"&institution=="domesticr1")))$citations, na.rm = TRUE)
muB <- mean(filter(df, ((lettergroup == "B and C"|lettergroup == "B Only"|lettergroup == "A and B")&role=="professor"&field=="math"&institution=="domesticr1"))$citations, na.rm = TRUE)
muC <- mean(filter(df, ((lettergroup == "C Only"|lettergroup == "B and C")&role=="professor"&field=="math"&institution=="domesticr1"))$citations, na.rm = TRUE)

val1 <- muA - muB
val2 <- muB - muC
val3 <- muA - muC

set.seed(0)
pop<- filter(df, ((role=="professor")&field=="math"&institution=="domesticr1"))

n_a <- length(na.omit(filter(df, (lettergroup == "A Only"|lettergroup == "A and B")&(role=="professor")&field=="math"&institution=="domesticr1")$citations))
n_b <- length(na.omit(filter(df, (lettergroup == "B and C"|lettergroup == "B Only"|lettergroup == "A and B")&(role=="professor")&field=="math"&institution=="domesticr1")$citations))
n_c <- length(na.omit(filter(df, (lettergroup == "C Only"|lettergroup == "B and C")&(role=="professor")&field=="math"&institution=="domesticr1")$citations))

dist1 <- meanPermutation(na.omit(filter(pop,(lettergroup == "A Only"|lettergroup == "A and B"|lettergroup == "B and C"|lettergroup == "B Only"))$citations),10000,n_a)
dist2 <- meanPermutation(na.omit(filter(pop,lettergroup == "B and C"|lettergroup == "B Only"|lettergroup == "A and B"|lettergroup == "C Only")$citations),10000,n_b)
dist3 <- meanPermutation(na.omit(filter(pop,lettergroup == "A Only"|lettergroup == "A and B"|lettergroup == "C Only"|lettergroup == "B and C")$citations),10000,n_c)


sum(dist1 < val1)/length(dist1)
sum(dist2 < val2)/length(dist2)
sum(dist3 < val3)/length(dist3)
```

## GS citperyear

```{r, echo = FALSE}
mean(filter(df, (lettergroup == "A Only"|lettergroup == "A and B"))$citperyear, na.rm = TRUE)
mean(filter(df, (lettergroup == "B and C"|lettergroup == "B Only"|lettergroup == "A and B"))$citperyear, na.rm = TRUE)
mean(filter(df, (lettergroup == "C Only"|lettergroup == "B and C"))$citperyear, na.rm = TRUE)
median(filter(df, (lettergroup == "A Only"|lettergroup == "A and B"))$citperyear, na.rm = TRUE)
median(filter(df, (lettergroup == "B and C"|lettergroup == "B Only"|lettergroup == "A and B"))$citperyear, na.rm = TRUE)
median(filter(df, (lettergroup == "C Only"|lettergroup == "B and C"))$citperyear, na.rm = TRUE)
```
```{r}
muA <- mean(filter(df, (lettergroup == "A Only"|lettergroup == "A and B"))$citperyear, na.rm = TRUE)
muB <- mean(filter(df, (lettergroup == "B and C"|lettergroup == "B Only"|lettergroup == "A and B"))$citperyear, na.rm = TRUE)
muC <- mean(filter(df, (lettergroup == "C Only"|lettergroup == "B and C"))$citperyear, na.rm = TRUE)

val1 <- muA - muB
val2 <- muB - muC
val3 <- muA - muC

set.seed(0)
pop<- df

n_a <- length(na.omit(filter(df, (lettergroup == "A Only"|lettergroup == "A and B"))$citperyear))
n_b <- length(na.omit(filter(df, (lettergroup == "B and C"|lettergroup == "B Only"|lettergroup == "A and B"))$citperyear))
n_c <- length(na.omit(filter(df, (lettergroup == "C Only"|lettergroup == "B and C"))$citperyear))

dist1 <- meanPermutation(na.omit(filter(pop,(lettergroup == "A Only"|lettergroup == "A and B"|lettergroup == "B and C"|lettergroup == "B Only"))$citperyear),10000,n_a)
dist2 <- meanPermutation(na.omit(filter(pop,lettergroup == "B and C"|lettergroup == "B Only"|lettergroup == "A and B"|lettergroup == "C Only")$citperyear),10000,n_b)
dist3 <- meanPermutation(na.omit(filter(pop,lettergroup == "A Only"|lettergroup == "A and B"|lettergroup == "C Only"|lettergroup == "B and C")$citperyear),10000,n_c)


sum(dist1 < val1)/length(dist1)
sum(dist2 < val2)/length(dist2)
sum(dist3 < val3)/length(dist3)
```

```{r, echo=FALSE}
sum(dist1 < val1)/length(dist1)
sum(dist2 < val2)/length(dist2)
sum(dist3 < val3)/length(dist3)
```

## GS citperyear r1 math professors only

```{r, echo = FALSE}
mean(filter(df, ((lettergroup == "A Only"|lettergroup == "A and B")&(role=="professor")&field=="math"&institution=="domesticr1"))$citperyear, na.rm = TRUE)
mean(filter(df, ((lettergroup == "B and C"|lettergroup == "B Only"|lettergroup == "A and B")&(role=="professor")&field=="math"&institution=="domesticr1"))$citperyear, na.rm = TRUE)
mean(filter(df, ((lettergroup == "C Only"|lettergroup == "B and C")&(role=="professor")&field=="math"&institution=="domesticr1"))$citperyear, na.rm = TRUE)
median(filter(df, ((lettergroup == "A Only"|lettergroup == "A and B")&(role=="professor")&field=="math"&institution=="domesticr1"))$citperyear, na.rm = TRUE)
median(filter(df, ((lettergroup == "B and C"|lettergroup == "B Only"|lettergroup == "A and B")&(role=="professor")&field=="math"&institution=="domesticr1"))$citperyear, na.rm = TRUE)
median(filter(df, ((lettergroup == "C Only"|lettergroup == "B and C")&(role=="professor")&field=="math"&institution=="domesticr1"))$citperyear, na.rm = TRUE)
```
```{r}
muA <- mean(filter(df, ((lettergroup == "A Only"|lettergroup == "A and B")&(role=="professor")&field=="math"&institution=="domesticr1"))$citperyear, na.rm = TRUE)
muB <- mean(filter(df, ((lettergroup == "B and C"|lettergroup == "B Only"|lettergroup == "A and B")&(role=="professor")&field=="math"&institution=="domesticr1"))$citperyear, na.rm = TRUE)
muC <- mean(filter(df, ((lettergroup == "C Only"|lettergroup == "B and C")&(role=="professor")&field=="math"&institution=="domesticr1"))$citperyear, na.rm = TRUE)

val1 <- muA - muB
val2 <- muB - muC
val3 <- muA - muC

set.seed(0)
pop<- filter(df, ((role=="professor")&field=="math"&institution=="domesticr1"))

n_a <- length(na.omit(filter(df, (lettergroup == "A Only"|lettergroup == "A and B")&institution=="domesticr1"&(role=="professor")&field=="math")$citperyear))
n_b <- length(na.omit(filter(df, (lettergroup == "B and C"|lettergroup == "B Only"|lettergroup == "A and B")&institution=="domesticr1"&(role=="professor")&field=="math")$citperyear))
n_c <- length(na.omit(filter(df, (lettergroup == "C Only"|lettergroup == "B and C")&institution=="domesticr1"&(role=="professor")&field=="math")$citperyear))

dist1 <- meanPermutation(na.omit(filter(pop,(lettergroup == "A Only"|lettergroup == "A and B"|lettergroup == "B and C"|lettergroup == "B Only"))$citperyear),10000,n_a)
dist2 <- meanPermutation(na.omit(filter(pop,lettergroup == "B and C"|lettergroup == "B Only"|lettergroup == "A and B"|lettergroup == "C Only")$citperyear),10000,n_b)
dist3 <- meanPermutation(na.omit(filter(pop,lettergroup == "A Only"|lettergroup == "A and B"|lettergroup == "C Only"|lettergroup == "B and C")$citperyear),10000,n_c)

sum(dist1 < val1)/length(dist1)
sum(dist2 < val2)/length(dist2)
sum(dist3 < val3)/length(dist3)
```

##Nans

```{r, echo=FALSE}
library(naniar)
vis_miss(df)
```
```{r, echo = FALSE}
restrictedData <- filter(df, (institution=="domesticr1"&role=="professor"&field=="math"))
vis_miss(restrictedData)
```

## Bin Matched Testing

```{r, echo=FALSE}
#with amscit
set.seed(0)
df3 <- filter(df, (age>0)&(amscit>0))

ageMatchedRandom <- function(){
  Xp <- filter(df3,lettergroup=="A Only"|lettergroup=="A and B")
  X <- Xp[sample(nrow(Xp),20),]
  Yp <- filter(df3,lettergroup=="B Only"|lettergroup=="A and B"|lettergroup=="B and C"|lettergroup=="C Only")
  #instantiate matrix
  Y <- matrix(0,nrow = 1, ncol = nrow(X))
  #for each person in subsample
  for (i in 1:nrow(X)){
    #get their age
    a <- X[i,]$age
    #create range
    low <- a-2
    high <- a+2
    #find all persons between low and high and then pick one randomly
    agematched <- filter(Yp, (age <= high & age >= low))$amscit
    random <- sample(agematched,1)
    #store
    Y[,i] <- random
  }
  return(mean(X$amscit)-mean(Y))
}

#perform test 1000 times
Dist <- matrix(0,nrow=1,ncol=1000)
for (i in 1:1000){
  Dist[,i]<-ageMatchedRandom()
}
hist(Dist, main="Age Matched Test for MathSciNet Citations")
```

```{r, echo=FALSE, include = FALSE}
sum(Dist>=0)
```


## Google Scholar and AMS

```{r}
cor(filter(df,df$citations>0 & df$amscit >0)$citations,filter(df,df$citations & df$amscit>0)$amscit)
plot(log(filter(df,df$citations>0 & df$amscit >0)$citations),log(filter(df,df$citations & df$amscit>0)$amscit))
```

## Distribution of Google Scholar Citations


We now apply a square root. 

```{r, echo=FALSE}
hist(df$citations^0.5,
     main = "Histogram of Google Scholar Citations^0.5",
     xlab = "Number of Citations")
```

This looks like an exponential distribution.

```{r, echo=FALSE}
Z <- na.omit(df$citations^0.5)
params <- as.list(fitdistr(Z, "exponential")$estimate)

qplot(sample = Z, geom = 'blank') +
  stat_qq(distribution = qexp, dparams = params)+
  stat_qq_line(distribution = qexp, dparams = params)+
  ggtitle("Exponential QQ-plot of Google Scholar Citations^0.5")
```

```{r, echo=FALSE}
hist(df$amscit^0.5,
     main = "Histogram of MathSciNet sqrt(citations)",
     xlab = "Number of Citations")
```


## Average Citations as a function of age
```{r, echo = FALSE}
# Initialize vectors to hold the mean citations of professors younger than a certain age.
x.youngest.age <- 15
x.oldest.age <-65
x.age.range <- x.youngest.age:x.oldest.age
year.span <- x.oldest.age - x.youngest.age
x.a <- vector(mode="numeric", length=year.span)
x.b <- vector(mode="numeric", length=year.span)
x.c <- vector(mode="numeric", length=year.span)

for (x.age in x.age.range)
{
  x.a[1+x.age - x.youngest.age] <- mean(filter(df, ((lettergroup == "A Only"|lettergroup == "A and B")&(role=="professor")&(age <= x.age)))$amscit, na.rm = TRUE)
  x.b[1+x.age - x.youngest.age] <- mean(filter(df, ((lettergroup == "B Only"|lettergroup == "A and B"|lettergroup == "B and C")&(role=="professor")&(age <= x.age)))$amscit, na.rm = TRUE)
  x.c[1+x.age - x.youngest.age] <- mean(filter(df, ((lettergroup == "C Only"|lettergroup == "B and C")&(role=="professor")&(age <= x.age)))$amscit, na.rm = TRUE)
}

plot(x.age.range,x.c,
  type="l",
  col="green",
  ylim=c(0,3000),
  main="Cumulative Mean MathSciNet Citations as a Function of PhD Age",
  ylab="Mean citations",
  xlab="PhD age",
  lwd=2)
lines(x.age.range,x.b,
  type="l",
  col="blue",
  lwd=2)
lines(x.age.range,x.a,
  type="l",
  col="red",
  lwd=2)
legend("topleft",
  c("Letter A","Letter B", "Letter C"),
  fill=c("red","blue", "green"))
```

```{r, echo = FALSE}
# Initialize vectors to hold the mean citations of professors younger than a certain age.
x.youngest.age <- 15
x.oldest.age <-65
x.age.range <- x.youngest.age:x.oldest.age
year.span <- x.oldest.age - x.youngest.age
x.a <- vector(mode="numeric", length=year.span)
x.b <- vector(mode="numeric", length=year.span)
x.c <- vector(mode="numeric", length=year.span)

for (x.age in x.age.range)
{
  x.a[1+x.age - x.youngest.age] <- median(filter(df, ((lettergroup == "A Only"|lettergroup == "A and B")&(role=="professor")&(age <= x.age)))$amscit, na.rm = TRUE)
  x.b[1+x.age - x.youngest.age] <- median(filter(df, ((lettergroup == "B Only"|lettergroup == "A and B"|lettergroup == "B and C")&(role=="professor")&(age <= x.age)))$amscit, na.rm = TRUE)
  x.c[1+x.age - x.youngest.age] <- median(filter(df, ((lettergroup == "C Only"|lettergroup == "B and C")&(role=="professor")&(age <= x.age)))$amscit, na.rm = TRUE)
}

plot(x.age.range,x.c,
  type="l",
  col="green",
  ylim=c(0,3000),
  main="Cumulative Median MathSciNet Citations as a Function of PhD Age",
  ylab="Median citations",
  xlab="PhD age",
  lwd=2)
lines(x.age.range,x.b,
  type="l",
  col="blue",
  lwd=2)
lines(x.age.range,x.a,
  type="l",
  col="red",
  lwd=2)
legend("topleft",
  c("Letter A","Letter B", "Letter C"),
  fill=c("red","blue", "green"))
```

## citperyear
```{r, echo = FALSE}
# Initialize vectors to hold the mean citations of professors younger than a certain age.
x.youngest.age <- 15
x.oldest.age <-65
x.age.range <- x.youngest.age:x.oldest.age
year.span <- x.oldest.age - x.youngest.age
x.a <- vector(mode="numeric", length=year.span)
x.b <- vector(mode="numeric", length=year.span)
x.c <- vector(mode="numeric", length=year.span)

for (x.age in x.age.range)
{
  x.a[1+x.age - x.youngest.age] <- mean(filter(df, ((lettergroup == "A Only"|lettergroup == "A and B")&(role=="professor")&(age <= x.age)))$amscitperyear, na.rm = TRUE)
  x.b[1+x.age - x.youngest.age] <- mean(filter(df, ((lettergroup == "B Only"|lettergroup == "A and B"|lettergroup == "B and C")&(role=="professor")&(age <= x.age)))$amscitperyear, na.rm = TRUE)
  x.c[1+x.age - x.youngest.age] <- mean(filter(df, ((lettergroup == "C Only"|lettergroup == "B and C")&(role=="professor")&(age <= x.age)))$amscitperyear, na.rm = TRUE)
}

plot(x.age.range,x.c,
  type="l",
  col="green",
  ylim=c(0,60),
  main="Cumulative Mean MathSciNet Cit/Year as a Function of PhD Age",
  ylab="Mean citations",
  xlab="PhD age",
  lwd=2)
lines(x.age.range,x.b,
  type="l",
  col="blue",
  lwd=2)
lines(x.age.range,x.a,
  type="l",
  col="red",
  lwd=2)
legend("topleft",
  c("Letter A","Letter B", "Letter C"),
  fill=c("red","blue", "green"))
```

```{r, echo = FALSE}
# Initialize vectors to hold the mean citations of professors younger than a certain age.
x.youngest.age <- 15
x.oldest.age <-65
x.age.range <- x.youngest.age:x.oldest.age
year.span <- x.oldest.age - x.youngest.age
x.a <- vector(mode="numeric", length=year.span)
x.b <- vector(mode="numeric", length=year.span)
x.c <- vector(mode="numeric", length=year.span)

for (x.age in x.age.range)
{
  x.a[1+x.age - x.youngest.age] <- mean(filter(df, ((lettergroup == "A Only"|lettergroup == "A and B")&(role=="professor")&(age <= x.age)))$amscitperyearpower, na.rm = TRUE)
  x.b[1+x.age - x.youngest.age] <- mean(filter(df, ((lettergroup == "B Only"|lettergroup == "A and B"|lettergroup == "B and C")&(role=="professor")&(age <= x.age)))$amscitperyearpower, na.rm = TRUE)
  x.c[1+x.age - x.youngest.age] <- mean(filter(df, ((lettergroup == "C Only"|lettergroup == "B and C")&(role=="professor")&(age <= x.age)))$amscitperyearpower, na.rm = TRUE)
}

plot(x.age.range,x.c,
  type="l",
  col="green",
  ylim=c(0,60),
  main="Cumulative Mean MathSciNet Cit/Year^1.3 as a Function of PhD Age",
  ylab="Mean citations",
  xlab="PhD age",
  lwd=2)
lines(x.age.range,x.b,
  type="l",
  col="blue",
  lwd=2)
lines(x.age.range,x.a,
  type="l",
  col="red",
  lwd=2)
legend("topleft",
  c("Letter A","Letter B", "Letter C"),
  fill=c("red","blue", "green"))
```


## age vs citations and citations per year

```{r, echo = FALSE}
pop <- filter(df, age>=0&amscit>=0)
fit <- lm(pop$amscit~pop$age)
fit
plot(pop$age,pop$amscit, main = "P1: Math Sci Net Citations vs Age", xlab = "Age", ylab="Citations")
abline(fit, col="red")
cf <- round(coef(fit), 1) 

## sign check to avoid having plus followed by minus for negative coefficients
eq <- paste0("Citations = ", cf[1],
             ifelse(sign(cf[2])==1, " + ", " - "), abs(cf[2]), "Age")

## printing of the equation
mtext(eq, 3, line=-2)
```

 the outlier is tao
 
 
```{r}
pop <- filter(df, age>=0&amscitperyear>=0)
fit <- lm(pop$amscitperyear~pop$age)
fit
plot(pop$age,pop$amscitperyear, main = "P2: Math Sci Net Citations per Year vs Age", xlab = "Age", ylab="Citations/Year")
abline(fit, col="red")
cf <- round(coef(fit), 2) 

## sign check to avoid having plus followed by minus for negative coefficients
eq <- paste0("Cit/Year = ", cf[1],
             ifelse(sign(cf[2])==1, " + ", " - "), abs(cf[2]), "Age")

## printing of the equation
mtext(eq, 3, line=-2)
confint(fit)
summary(fit)
```
```{r}
pop <- filter(df, age>=0&amscitperyearpower>=0)
fit <- lm(pop$amscitperyearpower~pop$age)
fit
plot(pop$age,pop$amscitperyearpower, main = "P3: Math Sci Net CitationsYear^1.3 vs Age", xlab = "Age", ylab="Citations/Year^1.3")
abline(fit, col="red")
cf <- round(coef(fit), 2) 

## sign check to avoid having plus followed by minus for negative coefficients
eq <- paste0("Cit/Year^1.3 = ", cf[1],
             ifelse(sign(cf[2])==1, " + ", " - "), abs(cf[2]), "Age")

## printing of the equation
mtext(eq, 3, line=-2)
confint(fit)
summary(fit)
```



## fellows

```{r}
length(filter(df,lettergroup=="A Only"|lettergroup=="A and B")$name)
```


```{r}
length(filter(df,lettergroup=="B Only"|lettergroup=="A and B"|lettergroup=="B and C")$name)
```

```{r}
length(filter(df,lettergroup=="C Only"|lettergroup=="B and C")$name)
```

